---
title: "Homework 8"
author: Robert LaBean
output: github_document
---


```{r, results='hide'}
library(tidyverse)
library(nycflights13)
library(babynames)
library(readxl)

```

Creating and accessing lists. 

```{r, tidy=TRUE}
mylist <- list(cars_df = mtcars, b_names_df = babynames, flights_df = flights)

glimpse(mylist$cars_df)
glimpse(mylist[2])
glimpse(mylist[["flights_df"]])

```

Access the mtcars dataframe.

```{r, tidy=TRUE}

glimpse(mtcars$mpg)
glimpse(mtcars[["mpg"]])

```

Adding "displacement per cylinder" column to mtcars.

```{r, tidy=TRUE}

# disp_cyl_mtcars <- mutate(mtcars, disp_cyl = mtcars$disp / mtcars$cyl)
# glimpse(disp_cyl_mtcars)

mtcars$disp_per_cyl = mtcars$disp / mtcars$cyl
glimpse(mtcars)


```

Working with factors. 

```{r, tidy=TRUE}

num.vec <- c(8, 6, 7, 5, 3, 0, 9)

as.factor(num.vec)

as.numeric(num.vec)

```
Load the BLS 2019 National Salary Data for all occupations, and since we're only interested in geoscientist data we will filter them out. 

```{r, tidy=TRUE}

salary_data <- read_xlsx("data/state_M2019_dl.xlsx") %>%
  filter(occ_code == "19-2042") %>%
  select(area_title, a_mean)
  glimpse(salary_data)
```
Since we want to compare the data state by state, we need to normalize by a cost of living index.
``` {r, tidy = TRUE}

col_index <- read.csv("data/col_index.csv", fileEncoding = "UTF-8-BOM")
norm_salary_data <- left_join(salary_data, col_index)
norm_salary_data$normalized = as.numeric(norm_salary_data$a_mean) / norm_salary_data$index
glimpse(norm_salary_data)

```

Now, let's plot it on a bar chart so it looks nice.

```{r}

salary_plot <- ggplot(norm_salary_data) + 
  geom_col(mapping = aes(area_title, normalized))+
  theme(axis.text.x = element_text(angle = 45))
print(salary_plot)