---
title: "Homework 09"
output: github_document
author: Robert LaBean
---

```{r message=FALSE}

library(tidyverse)

```

(1)

Following R for Data Science exercise 19.2.1.4, write a function to calculate the 
variance of a numeric vector. Be sure that the function auto-prints the result if
the result is not assigned to a variable.

```{r message=FALSE}

variancefunc <- function(x){
  
  sum(((x - mean(x)))^2)/(length(x)-1)
  
}

var_test <- sample(1:99999, 500)

```

Now let's test our function using the built in variance function.

``` {r message=FALSE}
  
vartest1 <- variancefunc(var_test)
vartestctrl <- var(var_test)

print(vartest1)
print(vartestctrl)

```

(2)

Write both_na(), a function that takes two vectors of the same length and returns 
the number of positions that have an NA in both vectors.

```{r message=FALSE}

both_na <- function(x, y){
  
  sum(is.na(x) & is.na(y))
  
}

```

Test the function

```{r message=FALSE}

na_1 <- c(1, 3, 5, NA, 7, NA, 8, 9, NA, NA, NA, 10, 19, NA)
na_2 <- c(NA, 1, 2, 3, NA, NA, NA, NA, NA, NA, NA, NA, 20, NA)

bnatest <- both_na(na_1, na_2)

print(bnatest)

```

(3)
Write a function to create a specific kind of plot, with useful labels, etc., 
that can take a data frame as an argument. This is useful when you want to make 
multiple, similar plots from different data.

```{r message=FALSE}

plot_df <- function(df, x, y, title, title_x, title_y){
  
ggplot(df)+
    geom_point(mapping = aes(x, y))+
    labs(title = title, x = title_x, y = title_y)
}

```

Test

```{r message=FALSE}

plot_df(mtcars, mtcars$mpg, mtcars$hp, "mpg vs hp", "mpg", "hp")

```
(4)

Write a function that takes one argument and does the following:

  if the argument is a numeric vector, multiplies every number in the vector by 2
  if the argument is a character vector, sorts the vector into alphabetical order
  if the argument is anything else, issues an informative warning and returns the 
  argument unchanged
  
```{r message=FALSE}

fun_vect <- function(x){
  
  for (i in x) {
    if (is.character(x)) {
    sort(x)
  }
    if (is.numeric(x)) {
      x*2
    }
    else (warning("Input is not a character or numeric vector"))
  }
}

```

Test

```{r message=FALSE}

fun_test <- sample(1:100, 75)
fun_test_char <- c("a", "b", "c", "d")

test1 <- fun_vect(fun_test)
test2 <- fun_vect(fun_test_char)

print(test1)
print(test2)

```

(5)

Write a function that tries to calculate the mean of a vector using mean(). If 
that is successful, it returns the mean of the vector. If mean() throws an error, 
it issues an informative message and returns the input intact.

```{r message=FALSE}

mean_test <- function(x){
  ifelse(is.numeric(x), mean(x), warning("X is not a numeric vector and the mean cannot be found"))
}

```

Test

```{r message=FALSE}

mean_test_vec1 <- sample(1:99999, 500)
mean_test_vec2 <- c(1, 3, "d", "e")

mtest <- mean_test(mean_test_vec1)
mtest2 <- mean_test(mean_test_vec2)

print(mtest)

```

(6)

Use system.time() (or, if you want to be more advanced about it, the bench or 
microbenchmark packages) to compare how long it takes to add two, 1-million-element
random vectors and store the result in a third vector:

 a. Using any vectorized operation (e.g., `+`)
 b. Using a for loop with a pre-allocated vector to store the results
 c. Using a for loop without pre-allocating a vector to store the results
 
```{r}

sum_time <- system.time(sample(1:10, 1000000, replace = TRUE) + 
                          sample(1:10, 1000000, replace = TRUE))

print(sum_time)

big_sum <- sample(1:10, 1000000, replace = TRUE)
big_sum2 <- sample(1:10, 1000000, replace = TRUE)

sum_time2 <- system.time(
  for (i in 1:1000000) {
    big_sum[i] + big_sum2[i]
  }
)
print(sum_time2)

# This needs to run overnight as it takes a VERY long time. 
# This takes entirely too long to run. Over 8 hours.
# Tested to 100,000 and it took ~20 minutes.

# sum_time3 <- system.time(
#   for (i in 1:1000000) {
#     sample(1:10, 1000000, replace = TRUE)[i] + sample(1:10, 1000000, replace = TRUE)[i] 
#   }
# )
# print(sum_time3)

```

