---
title: "Homework 5: Data Transformation, Exploration, and Plotting"
output: github_document
author: Robert LaBean
---


```{r message=FALSE}


library(tidyverse)
library(nycflights13)
library(dplyr)

```

Code to open documentation about flights.
```{r message=FALSE}

?nycflights13::flights

```

Using the glimpse function to look at the data inside flights.
```{r message=FALSE}

glimpse(flights)

```
There are `r nrow(flights)` rows in flights and `r ncol(flights)` columns.

Create a data-frame of only flights that departed 2 hours late or more.
```{r message=FALSE} 

late_flights <- filter(flights, dep_delay >= 120) 
percent_late_flights <-
  signif((nrow(late_flights)/nrow(flights))*100,3)
percent_late_flights # This is the percentage of flights that departed 2 hours late or later,
# rounded to 3 significant figures/


```
`r percent_late_flights`% of flights in 2013 were 2 hours late or more. 

Create a column that contains combined airline and flight number (e.g., "UA5401").
```{r message=FALSE}

airline_num <- mutate(flights, flightnum = paste0(flights$carrier, flights$flight))
airline_num # This column is carrier + flight number

```

Use `group_by()`, `summarise()`, and `n()` to make a data frame showing how many 
flights each aircraft (`tail_num()`) flew for each carrier throughout the entire data set. 
```{r}

num_flight_tot <- flights %>%
  group_by(carrier, tailnum)%>%
  summarise(numflights = n())
num_flight_tot # This column is the number of flights each tail number has flown for entire set of data