---
title: "Homework 10"
output: github_document
author: Robert LaBean
---
``` {r message = FALSE}

library(tidyverse)

```

(1)

Using the `diamonds` dataset, create a linear model of price as a function of the
weight of the diamond (`carat`). According to that model, what price would you 
expect for a 1-carat diamond?
```{r message=FALSE}

lm_diamonds <- lm(diamonds$price ~ diamonds$carat)
summary(lm_diamonds)

```

(2)

Now create a linear model of `price` as a function of `carat` and `color`.

``` {r message=FALSE}

lm_diamonds2 <- lm(diamonds$price ~ diamonds$carat + diamonds$color)
summary(lm_diamonds2)

# I'm doing two models to see what the difference in coercing diamonds$color 
# into a character vector first has. 

lm_diamonds3 <- lm(diamonds$price ~ diamonds$carat + as.character(diamonds$color))
summary(lm_diamonds2)

```

(3)

Make some plots of the raw data, and of the model fits, to make an argument as to which model is more useful.

```{r message=FALSE}
diam_plot_pc <- ggplot(diamonds) + 
  geom_point(mapping = aes(x = carat, y = price, colour = color)) +
  geom_smooth(method ='lm', mapping = aes(x = carat, y = price))

print(diam_plot_pc)

```

(4)

Write a function that returns some kind of structured result, which you define as
a new S3 class. Your class doesn’t need to be anything fancy, but it should be 
different from existing classes.

```{r message=FALSE}
good_couch <- function(
  is_comfy = TRUE, 
  my_favorite = "by window"
){

couch <- list(
  has_cushion = is_comfy, 
  favorite_spot = my_favorite
)

class(couch) <- append(class(couch), "good_couch")
return(couch)
}
my_couch <- good_couch()
my_couch

```

(5)

Write `print` and `summarise` methods for this class. These methods don’t need to be particularly useful, but they should be different from the default methods.

```{r message=FALSE}

print.good_couch <- function(x){
  cat("Print good_couch")
}
print(my_couch)

summarise.good_couch <- function(x){
  cat("summarise good_couch")
}

summarise(my_couch)
```